Insight Into Lumbar Back Pain \subtitle{What the Lumbar Spine Tells About Your Life}

### 2014-08-13 draft.tex
Started Results part

\subsection{Interactive Display of Decision Trees}
We have to create a decision tree for every non-image variable to see which one can be explained by image-derived variables.
%%
Since we have 134 non-image variables, the calculation yields a corresponding number of trees.
%%
Dividing the subjects into four groups using body-size, the number increases to 402 trees.
%%
To make the results cognitive feasible, we have to abstract the results of the classification.
%%------------
\subsubsection{Visualization of Classification Results}
%%
We follow the visual information-seeking mantra and want to provide an overview first, then details-on-demand \ref{shneiderman1996}.
%%
The optimal classification uses a few rules to precisely characterize the target variable.
%%
Therefore, we are interested in \emph{small trees} with a \emph{low classification error rate}.
%%
The error rate has to be judged depending on the number of classes of the target variable.
%%
An error rate of 50\% for a target variable with two manifestations makes the classifier equal to throwing a coin.
%%
The same error rate is acceptable a target variable with more than two manifestations.
%%
Picking a value for a variable of four manifestations with an error rate of 50\% is better than the random selection (error rate of 75\%).
%%
The following scales the error rate so that 0 denotes perfect classification and 1 is equal to random selection.
%%
\begin{equation}
error_{normalized} = \frac{error}{100 - \frac{1}{class number}}
\end{equation}
%%
This allows for comparability of error rates between variables.
%%
Abstracting the classification results into two continuos variables allows us to visualize the result of 134 classifications in one scatterplot.
%%
By mapping the group affiliation (in our case body-size) to color or shape, we can even plot all 402 results in a single visualization (Fig.~\ref{fig:decision-tree} a).
%%
\com{ToDo: Dummy-Variables}
%%------------
\subsubsection{Interaction With the Visualization}
%%
The described visualization provides a good overview over the classification results.
%%
We however still want to be able to look display \emph{details-on-demand} \ref{shneiderman1996} and examinate a decision tree in detail.
%%
This is realized by clicking on a entry on the visualization, which then displays the corresponding decision tree in detail (Fig.~\ref{fig:decision-tree} b).
%%
This allows for sequentially analyzing the interactions.
%%------------
\subsubsection{Implementation}
All analysis are carried out using \texttt{R}, a popular programming language for statistical calculations and visualizations.
%%
The interactive visualizations are realized using the \texttt{ggvis}\footnote{GGVIS Footnote} package.
%%
As opposed to the standard \texttt{R} plots, \texttt{ggvis} allows to adjust visualization parameters using UI controls, such as sliders.
%%
We implemented our application as web based prototype using \texttt{R Shiny}\footnote{R Shiny footnote}.
%%
The package combines the power of \texttt{R} with advantages of web-based applications.
%%
It allows us to quickly exchange results with our collaborating epidemiological partners.
%%
They can try out the technique without installing anything.
%%
Exchanging software becomes as easy as exchanging a hyperlink.
%%---------
\subsection{Results}

@@

\com{We can discriminate back pain using non-image data, but not with image-derived parameter. Welche parameter können wir gut unterscheiden}

>>>>

\subsection{Interactive Display of Decision Trees}
We have to create a decision tree for every non-image variable to see which one can be explained by image-derived variables.
%%
Since we have 134 non-image variables, the calculation yields a corresponding number of trees.
%%
Dividing the subjects into four groups using body-size, the number increases to 402 trees.
%%
To make the results cognitive feasible, we have to abstract the results of the classification.
%%------------
\subsubsection{Visualization of Classification Results}
%%
We follow the visual information-seeking mantra and want to provide an overview first, then details-on-demand \ref{shneiderman1996}.
%%
The optimal classification uses a few rules to precisely characterize the target variable.
%%
Therefore, we are interested in \emph{small trees} with a \emph{low classification error rate}.
%%
The error rate has to be judged depending on the number of classes of the target variable.
%%
An error rate of 50\% for a target variable with two manifestations makes the classifier equal to throwing a coin.
%%
The same error rate is acceptable a target variable with more than two manifestations.
%%
Picking a value for a variable of four manifestations with an error rate of 50\% is better than the random selection (error rate of 75\%).
%%
The following scales the error rate so that 0 denotes perfect classification and 1 is equal to random selection.
%%
\begin{equation}
error_{normalized} = \frac{error}{100 - \frac{100}{class number}}
\end{equation}
%%
This allows for comparability of error rates between variables.
%%
Abstracting the classification results into two continuos variables allows us to visualize the result of 134 classifications in one scatterplot.
%%
By mapping the group affiliation (in our case body-size) to color or shape, we can even plot all 402 results in a single visualization (Fig.~\ref{fig:decision-tree} a).
%%

Many follow-up variables are sparse, such as medication of diabetes or reason of early retirement.
%%
The classification algorithm may produce higher accuracy for variables with less subjects due to the small sample size.
%%
This makes these results less reliable.
%%
We map the number of subjects associated with a variable to point diameter in the scatter plot.
%%
This allows to instantly assess the reliability of the result.
%%
\subsubsection{Dummy Variables}
Dummy variables abstract a categorical variable with multiple manifestations into several binary variables.
%%
Each binary variables encodes the presence of a manifestation.
%%
For example, a pain indicator variable ranging from \emph{1 - no pain} to \emph{4 - large pain} is subdivided into four binary variables (\emph{No pain - Yes/No} to \emph{Large Pain - Yes/no}).
%%
One subject can only have one of these variables set to true.
%%
This is useful for our classification, because it allows to determine which manifestation of a variable can be described best using the image data parameter.
%%\com{ToDo: Dummy-Variables}
%%------------
\subsubsection{Interaction With the Visualization}
%%
The described visualization provides a good overview over the classification results.
%%
We however still want to be able to look display \emph{details-on-demand} \ref{shneiderman1996} and examine a decision tree in detail.
%%
This is realized by clicking on a entry on the visualization, which then displays the corresponding decision tree in detail (Fig.~\ref{fig:decision-tree} b).
%%
This allows for sequentially analyzing the interactions.
%%------------
\subsubsection{Implementation}
All analysis are carried out using \texttt{R}, a popular programming language for statistical calculations and visualizations.
%%
The interactive visualizations are realized using the \texttt{ggvis}\footnote{GGVIS Footnote} package.
%%
As opposed to the standard \texttt{R} plots, \texttt{ggvis} allows to adjust visualization parameters using UI controls, such as sliders.
%%
We implemented our application as web based prototype using \texttt{R Shiny}\footnote{R Shiny footnote}.
%%
The package combines the power of \texttt{R} with advantages of web-based applications.
%%
It allows us to quickly exchange results with our collaborating epidemiological partners.
%%
They can try out the technique without installing anything.
%%
Exchanging software becomes as easy as exchanging a hyperlink.
%%---------
\subsection{Results}
We ran the analysis using different subject groups:
\begin{enumerate}
	%%
	\item All subjects,
	%%
	\item subdivision into \emph{male} and \emph{female},
	%%
	\item subdivision into \emph{age quantiles}, yielding the groups (21, 43.8] (43.8, 54] (54, 64] (64, 90],
	%%
	\item subdivision into \emph{size quantiles}, yielding the groups (139, 164], (164, 171], (171, 177], (177, 202].
	%%
\end{enumerate}

\subsubsection{All Variables}
% None of the pain indicators can be reliably described using the image-based parameter.
None of the pain indicators can be described reliably using the image-based parameter.
%%
The perfect categorization of the \emph{dialysis} parameter was conspicuous.
%%
This is due to the subject distribution among the manifestation.
%%
Only two subjects receive a dialysis.
%%
The good discrimination is not meaningful.
%%
Gender can be discriminated with an error rate of 31,5\% using 7 rules and incorporates only curvature and curvature angle variables.
%%
Surprisingly high is the classification of medication for high blood pressure.
%%
1,058 subjects are classified with an error rate of 23,8\% solely based on coronal mean curvature.
%%
Other features well classified include marital status, foot pain and number of born children.
%%

The dummy variable analysis yields fewer results.
%%
It shows that subjects between 64 and 90 years of age can be detected with 23\% error rate by using the curvature angle and coronal mean curvature.
%%
Subjects with body size between 177 cm and 202 cm can classified with an error rate of 22,1\%. solely using mean curvature.
%%
\subsubsection{Gender Groups}
\subsubsection{Age Groups}
\subsubsection{Size Groups}
%%
Many previously described results can be associated to be confounded with subject size.
%%
Differences in the gender analysis are mostly due to the average height difference between males and females.
%%
For example, large subjects are already characterized by their rather straight spine.
%
The question is, whether the inter-group spine-variability parameter is enough for predicting other parameter or not.
%%
Dividing subjects into height groups potentially highlights classifications not confounded by body-height.
%%

@@

\com{We can discriminate back pain using non-image data, but not with image-derived parameter. Welche parameter koennen wir gut unterscheiden}

\com{Allgmeine Analyse der Daten mit einbringen. Hierfuer statistisch Auswertung machen, welche Parameter am meisten verwendet wurden bei der Kategorisierung. Nur Parameter verwenden, die genauer sind als 0.5\%, der Entscheidungsbaum kleiner als 15 ist und das Feature mindestens 100 Probanden enthält }
